In this file, I will apply some examples regarding IRT, then will use the codes later. 
```{r}
library("MPsychoR")
library("Gifi")
library("mirt")
library("tidyverse")
data("zareki")
zarsub <- zareki[, grep("subtr", colnames(zareki))]
```

```{r}
install.packages("Gifi")
```

#Testing the dimensiality of the scale using a plot.. The arrows indicates where the questions are. 

```{r}
prinzar <- princals(zarsub)
plot(prinzar, main = "Zareki Loadings")
```


Changing data from long to wide


<<<<<<< HEAD
=======

>>>>>>> a6a03a20dee958bf35b5a42be891c6baa0d54f25
```{r}
PreCCscores <- read.csv("PreCCscores_Complete.csv")
```


```{r}
PreCCscores <- select(PreCCscores,ID, Question, value)
```

<<<<<<< HEAD
```{r}
trial <- pivot_wider(PreCCscores, names_from = Question, values_from = value, values_fn = NULL)
```

```{r}
as.data.frame(trial)
```

```{r}
head(trial)
```

=======

```{r}
pivot_wider(PreCCscores, names_from = Question, values_from = value)
```


```{r}
PreCCscores  %>%
pivot_wider_spec(names_repair = "check_unique", names_from = Question, values_from = value)
```
>>>>>>> a6a03a20dee958bf35b5a42be891c6baa0d54f25
Duplicated items

```{r}
spread(PreCCscores, Question, value)
```

<<<<<<< HEAD
same there are some duplicated values 
```{r}
PreCC_unique: 
```

```{r}
PreCCPrincals <- princals(trialmatrix)
plot(PreCCPrincals, main = "Density Questionnaire Loadings")
```


Selecting the dichotomous variables only. 
```{r}
trialselect <- select(trial, -Q09, -Q14, -Q20, -Q21, -Q22)
```


```{r}
PreCCPrincals <- princals(trial2, ordinal = TRUE, active = "Q08")
=======
```{r}
duplicated(PreCCscores$ID)
```

```{r}

```

```{r}
PreCCPrincals <- princals(PreCCscores)
>>>>>>> a6a03a20dee958bf35b5a42be891c6baa0d54f25
plot(PreCCPrincals, main = "Density Questionnaire Loadings")
```

