---
title: "PALS"
author: "Tamer Said"
date: "01/11/2020"
output: html_document
---

Summary of the PALS Questions and their factors
Mastery Goals (MG): Q01:Q05
Performance Approach Goals (PAPG) Q06:Q10
Performance Avoidance Goals (PAVG) Q11:Q14
Self-efficacy (SE) Q15:Q19

```{r}
PreQ <- read.csv("PreQ_filtered.csv")
```

```{r}
library(tidyverse)
```
Selecting  the columns in PALS

```{r}
PALS <- select(.data = PreQ,ID,Gender, Certificate, Q27_1:Q27_19)
```

#Renaming question numbers

```{r}
PALS <- rename(PALS, Q01 = Q27_1, Q02 = Q27_2, Q03= Q27_3, Q04= Q27_4, Q05 = Q27_5, Q06 = Q27_6, Q07= Q27_7, Q08= Q27_8, Q09= Q27_9, Q10=Q27_10, Q11=Q27_11, Q12=Q27_12, Q13=Q27_13, Q14 = Q27_14, Q15= Q27_15, Q16=Q27_16, Q17=Q27_17, Q18=Q27_18, Q19=Q27_19)

```

Saving PALS into csv file

```{r}
write.csv(PALS, "PALS-R.csv")
```

```{r}
PALS <- read.csv("PALS-mod.csv")
```
I saved an earlier version of this file, and loaded it here again.

Now I am going to arrange the PALS responses in long format 

```{r}
PALS_long <- pivot_longer(data = PALS,cols =  Q01:Q19,names_to = "Question",values_to = "Response"  )
```
It appears that the long data format is affecting my analysis, since each case is counted 15 times. 

Loading scores template

```{r}
PALS_scoring <- read.csv("Scoring_PALS.csv")
```

Joining the two tables to calculate scores

```{r}
PALS_joined <- inner_join (x = PALS,y= PALS_scoring, by = "Response", group_by(ID))
```

Calculating scores
```{r}
PALS_scores <- PALS_joined %>% 
             group_by(ID)
             
```


```{r}
PALS_scores <- PALS_joined %>% 
  group_by(ID) %>%
    summarise (PALS_Score = sum(score))
```

Saving ID numbers and PALS scores

```{r}
write.csv(PALS_scores, "PALS_scores.csv")
```

#Initial analysis before CFA.

I now need to group students' responses by question to do my analysis.

```{r}
PALS_Q <- PALS_joined %>% 
  group_by(Question)
```


```{r}

```

Loading #libraries 

```{r}
library(tidyverse) # For data wrangling
library(ggplot2) # for graphing / figures - mostly histograms/ppplots here
library(pastecs) # Needed to run normality tests
library(car) # Needed to run Levene test
library(lsr) # Navarro package for running psychology tests
library(psych) # for key psychology stats
library(effects) # Effects package, needed for the estimated means, includes lower/upper 95% conf limits
options(scipen=99) # This is to indicate how many digits after the decimal, this one is for 2 digits, but can be changed
```


I need to create a total coloumn for the score of each participant in the PALS questions
First,I will create total variable to sum numeric values across rows:

```{r}
total = rowSums(Filter(is.numeric, PALS), na.rm = TRUE)
```

#Calcualting total score for each participant. 
Now, I will create a sum coloumn for the total of student's responses in the questions


```{r}
PALS <- mutate(.data = PALS,sum = total)
```
#clearing missing values
It appeared that soem students did not attempt the survey at all, or filled only one question. The total of that is five students, in rows 13, 451, 75, 401, 192. 
Instead of typing every row number, I would select those who scored more than 6, i.e. those who have completed the survey.
```{r}
PALSNM <- subset(PALS, PALS$sum > 6, select = ID:sum)
```
PALSNM is PALS no missing values. 


```{r}
write.csv(PALS, "PALS.csv")
```
#Parametric assumptions

This can be done via histograms on the sum of the scores. I am not sure of shall I do this on the responses for each question?

```{r}
PALS_hist <- ggplot(PALS, aes(sum)) + 
geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
labs(x = "score", y = "")
```
Adding normal curve

```{r}
PALS_hist + stat_function(fun = dnorm, args = list(mean = mean(PALS$sum,
na.rm = TRUE), sd = sd(PALS_Q$score, na.rm = TRUE)), colour = "black", size = 1)
```

Drawing Q-Plot to check the distribution of my scores compared to normal distribution

```{r}
qqplot.scores <- qplot(sample = PALS_Q$score, stat="qq")
```
There was an error with stat function and it says it is deprecated, I am not sure whether the resulting qplot is meanigful or not. 
 
Descrbing the scores of per student

```{r}
describe(PALS$sum)
```

Shapiro test:
```{r}
round (stat.desc(PALS$sum, basic = FALSE, norm = TRUE), digits= 3)
```
#Notes on normality of scores:
skewness: -ve score indicate a build up of high scores
kurtosis: 3.173 : Positive values of kurtosis indicate a pointy and heavy-tailed distribution.

-if skew.2SE (-3.4) or kurt.2SE(7.6) are greater than 1 (ignoring the plus or minus sign) then you have
significant skew/kurtosis (at p < .05)
```{r}
shapiro.test(PALS$sum)
```

#Shapiro test:
The shapiro scores is W = 0.95313, p-value = 0.000000000003767, which shows a significant non-normality. However, since the sample size is large, one shall be careful when interpreting these results since they get distorted at large sample sizes (Field, 2002)

Shapiro test across genders

```{r}
by(PALS$sum, PALS$Gender, shapiro.test)
```

Descriptive stats split by gender

```{r}
describeBy(PALS$sum, group = PALS$Gender)
```

```{r}
describeBy(PALS$sum, group = "Question")
```


Loading ggpubr library for graphs, then creating QQ plot to compare the deviation of my sample from the theoretical nornmal distribution

```{r}
library(ggpubr)
ggqqplot(PALS$sum)
```
The graph showed a normal distribution  when compared to the theoritical, it appears that there are some outliers, probably those who did not fill the PALS. 

Tried to do a dot chart for each question to compare their scores. 

```{r}
ggdotchart(PALS_Q$score, x = "question", y = "density",
           color = "Question",                                # Color by groups
           palette = c("#00AFBB", "#E7B800", "#FC4E07"), # Custom color palette
           sorting = "ascending",                        # Sort value in descending order
           add = "segments",                             # Add segments from y = 0 to dots
           ggtheme = theme_pubr()                        # ggplot2 theme
           )
```
Code did not work

```{r}
ggbarplot(PALS_Q, x = "Question", y = "Response",
          fill = "Question", sort.by.groups = TRUE)
```
The code worked, but the color of the bars did not show up.

Trying a dotchart
```{r}
ggdotchart(PALS_Q, x = "Question", y = "score",
           color = "Question",                                 # Custom color palette
           sorting = "descending",                       # Sort value in descending order
                                                   # Rotate vertically
           group = "Question",                                # Order by groups
           dot.size = 6,                                 # Large dot size
           label = round(PALS_Q$score),                        # Add mpg values as dot labels
           font.label = list(color = "white", size = 9,
                             vjust = 0.5),               # Adjust label parameters
           ggtheme = theme_pubr()                        # ggplot2 theme
           )
```
```{r}

```


#Homeogenity of variance

Using leven's test:
leveneTest(outcome variable, group, center = median/mean)

Testing the variance across gender.

```{r}
leveneTest(PALS$sum, PALS$Gender)
```
Since pr is 0.34, it means that the variance across gender is not signigicanmt
Next,I need to run the test for each question, to check the variance across questions.  

```{r}
leveneTest(PALS_Q$score, PALS_Q$Question)
```
F(1,10254)=45.3, p= 0.02, this result shows a significant variance across the questions. This result should be interpreted with caution, since large sample sizes may make Levene's test signficiant, even if the variance is not great.
Another test that is used is the variance ration (not sure though if need)

```{r}

```

#Correlations

```{r}
install.packages("Hmisc"); install.packages("ggm");
install.packages("ggplot2"); install.packages("polycor")
```

```{r}
library(boot); library(ggm); library(ggplot2); library(Hmisc);
library(polycor)
```
Running correlations for the whole data frame. This command below excludes the missing values and those that are not numeric.
```{r}
cor(PALS[sapply(PALS, is.numeric)], use='pairwise')
```
Saving the correlation matrix to data

```{r}
PALScorr <- cor(PALS[sapply(PALS, is.numeric)], use='pairwise')
```


Running Barlette test on the correlation matrix in prep for FA. The p value shall be less than 0.5
A significant test tells us that the R-matrix is not an identity matrix(identity matrix is the case when variables are NOT related to each other and corr is zero); Therefore, if it is significant
then it means that the correlations between variables are (overall) significantly different
from zero, if Bartlett’s test is significant then it is good news (Field, 2009)


```{r}
cortest.bartlett(PALScorr, n=541)
```

Running KMO test for the degree of common variance (ideally, a score of 0.5 or higher is good). values close to 1 indicates that patterns of correlations are relatively compact and so factor analysis should yield distinct and reliable factors.

```{r}
KMO(PALScorr)
```

Finding the #determinant

```{r}
det(PALScorr)
```

```{r}
cor(PALS$sum, PALS$Gender, use = "complete.obs", method = "pearson")
```

It did not work, since Gender is a class variable and it needs to be numeric for correlation to take place. 

Some correlation commands do not work on data frames, so would have to convert my df into matrix.

```{r}
PALSMatrix <- as.matrix(PALSNM)
```
Runnning general correlation on the whole df, and saving it into PALScorr

```{r}
PALScorr <- cor(PALS[sapply(PALS, is.numeric)], use='pairwise')
```


#EFA: I need to start with EFA first and some preparatory analysis

Installing packages and loading libraries
```{r}
install.packages("corpcor"); install.packages("GPArotation"); install.
packages("psych")

library(corpcor); library(GPArotation); library(psych)
```

#PCA:I have set the factors to 19, which is the number of the items. This is just an exploratory stem

```{r}
pc19 <- principal(PALScorr, nfactors = 19, rotate = "none")
```

```{r}
pc19
```
SS loading: refer to sum of squared loadings
```{r}
pc19$values

```

This command allows to view the eigenvalues of each factor. 

It appears that factor one 6.29 units of the variance in the model ,which = 6.29/19 = 0.33 or 33% of the variance 
Screeplot to determine the eigenvalues
```{r}
plot(pc19$values, type = "b")
```
Based on the scree plot, it may be feasible to test the 4 components model, since the inflection point is in the 4th component. This also agres with my theoritical model.

```{r}
pc04 <- principal(PALScorr, nfactors = 4, rotate = "none")
pc04
```
#h2 (commonalities) show the amount of variance in each variable that can
be explained by the retained factors is represented by the communalities after extraction

we can also obtain the factor model using this function

```{r}
factor.model(pc04$loadings)
```

To compare our factor loadings with the orginial corr. matrix, we use the following:

```{r}
factor.residuals(PALScorr, pc04$loadings)
```
The above output  contains the differences between the observed correlation
coefficients and the ones predicted from the model. For a good model these values will all be small.

Saving the residuals in a separate object

```{r}
residuals <- factor.residuals(PALScorr, pc04$loadings)
```

To extract the upper triangle of the residuals and save it as a matrix

```{r}
residuals<-as.matrix(residuals[upper.tri(residuals)])
```
i will save the large residuals (> 0.05) in a separate object

```{r}
large.resid<-abs(residuals) > 0.05
```

The residuals that are greater than 0.05 will be labelled as True, others will be False.

```{r}
sum(large.resid)
```
The sum is 43, to see the proportion of these,
```{r}
sum(large.resid)/nrow(residuals)
```

This means the 22% of our data have residuals larger than 0.05. Which is a good indicator

To view the residuals graphically

```{r}
hist(residuals)
```
As we can see, most of the residuals fall below 0.05

After doing EFA, let's move to CFA

Loading libraries

```{r}
library(tidyverse) # For data wrangling
library(lavaan) # For CFA/MI/SEM
library(semPlot) # For CFA/MI/SEM
library (semTools) # For CFA/MI/SEM
library(OpenMx) # For SEM
library(pastecs) # Needed to run normatlity tests
library(car) # Needed to run Levene test
library(lsr) # Navarro package for running psychology tests
library(psych) # for key psychology stats
library(effects) # Effects package, needed for the estimated means, includses lower/upper 95% conf limits
options(scipen=99) # This is to indicate how many digits after the decimal, this one is for 2 digits, but can be changed
```

# CFA and 1-defining the model

Mastery Goals (MG): Q01:Q05
Performance Approach Goals (PAPG) Q06:Q10
Performance Avoidance Goals (PAVG) Q11:Q14
Self-efficacy (SE) Q15:Q19
Let's start be defining the measurement model

```{r}
latents.model <- 'MasteryGoals =~ Q01 + Q02 + Q03 + Q04 + Q05
PerfAppGoals =~ Q06 + Q07 + Q08 + Q09 + Q10
PerfAvGoals =~ Q11 + Q12 + Q13 + Q14
SelfEfficacy =~ Q15 + Q16 + Q17 + Q18 + Q19'
```

Run CFA model


```{r}
PALSfit1 <- cfa(latents.model, data = PALS)#stardard fit
summary(PALSfit) #to view summary of the fit
partable(PALSfit) # to have the summary of parameters converted to a dataframe
vartable(PALSfit) # to have the summary of variables converted to a dataframe
parameterestimates(PALSfit) #parameterEstimates() includes the estimates, standard errors, z value, p value, and 95% CIs for all model parameters.
```
To see a summary of the tests and esimtates of the model

```{r}
summary(PALSfit,fit.measures = TRUE )
```
#Notes on the fit measures
CFI and TLI needs to be high, as they compare your model to a completely free model. Comparative Fit Index (CFI)                    0.920
Tucker-Lewis Index (TLI)                       0.907 -
RMSEA: needs to be low as possible

plotting a sem diagram

```{r}
semPaths(PALSfit, title= FALSE, curvePivot= TRUE)
```
Another view of viewing the plot

```{r}
semPaths(PALSfit1, "std", edge.label.cex=0.5, 
         curvePivot = TRUE, exoVar = FALSE)
```

Note: the thicker the lines, the higher the loadings. 

Will now try Michelle's script and save it as fit2

```{r}
PALSfit2 <- cfa(latents.model, data = PALS, se = "robust", test = "Satorra-Bentler") 
summary(PALSfit, rsq = TRUE,fit.measures = TRUE)
fitmeasures(PALSfit)
varTable(PALSfit)
parameterEstimates(PALSfit)
```
Notes on the #Fitness of the model:

Root Mean Square Error of Approximation (RMSEA): a perfect fit will result in a RMSEA = 0, the closer the no. to zer0, the better the fit. Here my RMSEA = 0.067. Based on Hu and Bentler (1999), an RMSEA value of 0.06 ir bewlow is recommended. MacCallum (1996) described RMSEA values between 0.08-0.10 to be 'mediocre'. 

Comparative Fit index (CFI): compares the output of my model to a totally constrained model, the range of values is from 0-1, the closer the value to 1, the better my model is. My CFI score is: 0.92. Based on Hu and Bentler (1999), a CFI value of 0.95 or greater is recommended. Bentler was more lenient by stating that CFI & TLI values between 0.9 and 0.95 may indicated an acceptable model fit. 

Tucker-Lewis Index (TLI): it adds penalty to the free parameters that do not markedly improve the fit of the model. My TLI score is: 0.907. 



Comparing the fits 

```{r}
compareFit(PALSfit1, PALSfit2, nested=FALSE)
```
Exactly the same outcome..it seems that there is no differnece between the two methods. (check with Michelle)

More plots (from Michelle)

```{r}
 #the model plot below shows up as a black & white plot with nonstandardised parameter estimates
semPaths(PALSfit, what = "path", whatLabels = "par", style = "lisrel", layout = "tree", 
         intercepts = TRUE, residuals = TRUE, thresholds = TRUE, 
         nCharNodes = 8, nCharEdges = 3, 
         sizeMan = 9, sizeMan2 = 4, sizeLat = 9, sizeLat2 = 7, sizeInt = 6, sizeInt2 = 4)
# the model plot below shows up as a black & white plot with standardised parameter estimates
semPaths(PALSfit, what = "path", whatLabels = "std", style = "lisrel", layout = "tree", 
         intercepts = TRUE, residuals = TRUE, thresholds = TRUE, 
         nCharNodes = 8, nCharEdges = 3, 
         sizeMan = 9, sizeMan2 = 4, sizeLat = 9, sizeLat2 = 7, sizeInt = 6, sizeInt2 = 4)
```
# Standardized vs. non-standardized parameters:

The above plot, we can see 2 ways of measuring the scale of the latents, in the non-standardized, the first factor is set to 1.0, which is also called the marker variable. This is used as a reference for the other variables in the model. The other way to measure the scale of the latents, is to standardize the scores (using Z-scores)



