---
title: "Data Cleaning"
author: "Tamer Said"
date: "27/10/2020"
output: html_document
---

This file is for cleaning my pre-q data, to clean duplicates and change the wrong names inserted in the ID

```{r}
Pre-Q <- read.table("pre-quest-unique.csv")
```

```{r}
PreQ<-read.csv("pre-quest-unique.csv", header = TRUE)

```

```{r}
library(tidyverse)
```
Selecting only the variables of interest from the dataset

```{r}
PreQ_filtered <- select(PreQ, ID, StartDate, Finished,Q1:SC0)
```

Selecting only those who have finished the survey

```{r}
PreQ_filtered <- filter(PreQ_filtered, Finished == "TRUE")
```

Retained 563 obs of 607, who have finished the survey. 


```{r}
PreQ_filtered<- filter(PreQ_filtered,Q1>1)
```

After retaining only those who have IDs more than 1, I am left with 556 obs from 563 observations.

#Renaming variables



#Changing some ID numbers

One student entered a wrong ID number 422 instead of 4229

```{r}
Demog$ID[Demog$ID=="422"] <- "4229"
```

replacing 1657 with 4657

```{r}
Demog$ID[Demog$ID=="1657"] <- "4657"
```
 Replacing 1499 (bear strawberry) with 4799
 
```{r}
Demog$ID[Demog$ID=="1499"] <- "4799"
```
 
I noitced that rows 412 and 413 have the same DoB, but there was a cofnsution in the ID no. (2008 & 2018) Since he did not enter the picture or ID number,I will have to delete his entry.. or can I match it using other credentials in the post-survey?

Another student entered 3863 instead of 4863, so I will have to change this 
```{r}
Demog$ID[Demog$ID=="3863"] <- "4863"
```

#Another student entered 4036 instead of 4952

```{r}
Demog$ID[Demog$ID=="4952"] <- "4036"
```

#A wrong ID of 43337 (green bat) was entered instead of 4337

```{r}
Demog$ID[Demog$ID=="43337"] <- "4337"
```

#A wrong value of 9424 (gold balloon) placed instead of 4924


this code is better to use, since I can use it in another files to be consistent with the changes.. 
```{r}
Demog$ID[Demog$ID=="9424"] <- "4924"
```


